---
title: "Getting started with brochure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with brochure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

A `{shiny}` app usually lives at a single url, and everything you show is a
matter of hiding and showing pieces of one page. `{brochure}` takes the other
road: an app serves several endpoints, and each of them is its own page, with
its own UI and its own server function.

```{r}
library(shiny)
library(brochure)
```

## A page

`page()` binds an href to a UI and a server function. The server is optional:
a page that only shows static content does not need one.

```{r}
home <- function() {
  page(
    href = "/",
    ui = tagList(
      h1("Home"),
      plotOutput("plot")
    ),
    server = function(input, output, session) {
      output$plot <- renderPlot(plot(mtcars))
    }
  )
}

contact <- function() {
  page(
    href = "/contact",
    ui = tagList(
      h1("Contact"),
      tags$p("here@example.com")
    )
  )
}
```

## An app

`brochureApp()` replaces `shinyApp()`, and takes the pages:

```{r}
brochureApp(
  home(),
  contact()
)
```

Navigating between pages is done with plain links — `tags$a(href = "/contact")`.
There is no router to configure on the client: a link is a real navigation, and
the browser asks the server for that url.

Which brings the one thing to keep in mind: **every page is a new Shiny
session**. Nothing you compute on `/` is available on `/contact`. That is the
point rather than a limitation — it forces the data flowing between pages to be
explicit — but it means state has to live somewhere shared: a cookie, a
database, a disk cache.

## Parameters in the url

An href can carry parameters, `{routr}` style. Read them back with
`get_keys()`: with no argument inside the server, and with the `request` the ui
receives inside the ui.

```{r}
brochureApp(
  page(
    href = "/who/:id",
    ui = function(request) {
      tagList(
        h1(sprintf("Hello %s", get_keys(request)$id)),
        verbatimTextOutput("from_server")
      )
    },
    server = function(input, output, session) {
      output$from_server <- renderText(get_keys()$id)
    }
  )
)
```

`/who/colin` and `/who/fay` are two sessions of the same page, each with its
own `id`.

## Redirections

```{r}
brochureApp(
  home(),
  redirect(from = "/index.html", to = "/")
)
```

`redirect()` answers at the http level, before any Shiny code runs. From inside
a server, `server_redirect("/contact")` sends the browser to another page.

## Middleware

Every page, and the app itself, takes `req_handlers` and `res_handlers`: lists
of functions run on the request before the response is built, and on the
response before it is sent. App level handlers run first, then those of the
matched page.

A `req_handler` returning an `httpResponse` short-circuits everything else,
which is how you serve something that is not a page at all:

```{r}
brochureApp(
  home(),
  page(
    href = "/healthcheck",
    ui = tagList(),
    req_handlers = list(
      ~ shiny::httpResponse(200, content = "OK")
    )
  )
)
```

A `res_handler` is where you set a cookie:

```{r}
page(
  href = "/login",
  ui = tagList(h1("Logged in")),
  res_handlers = list(
    ~ set_cookie(.x, "SESSION", "abc123")
  )
)
```

Read it back from the server with `get_cookies()` and `parse_cookie_string()`.

## Behind a reverse proxy

If the app is not served at the root of its domain, tell it where it is
mounted with `basepath`:

```{r}
brochureApp(
  home(),
  basepath = "myapp"
)
```

The prefix is removed from incoming urls, so it keeps matching your `page()`
hrefs, and prepended to the urls the app emits. On Posit Connect the mount is
picked up on its own and `basepath` is not needed.

## What's next

- `vignette("handlers")` — running code before and after a page is built,
  and answering a request without Shiny at all.
- `vignette("cookies")` — setting and reading cookies, and carrying a session
  from one page to the next.
- `vignette("deployment")` — serving the app under a prefix.

The README covers the golem integration and the design patterns for sharing
state between pages.
