---
title: "Cookies and state between pages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookies and state between pages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r}
library(shiny)
library(brochure)
```

Every page of a brochure app is a new Shiny session. Nothing computed on one
page survives a navigation to another. That is deliberate, and it means state
shared between pages has to be written down somewhere: a cookie, and usually a
backend keyed by what the cookie holds.

## Setting a cookie

A cookie is a response header, so it is set from a `res_handler`:

```{r}
page(
  href = "/login",
  ui = tagList(h1("You are logged in")),
  res_handlers = list(
    ~ set_cookie(.x, "SESSION", "a-random-id", path = "/")
  )
)
```

`set_cookie()` defaults to `http_only = TRUE` and `same_site = "Lax"`, so the
cookie is not readable from javascript and is not sent on cross-site requests.
Pass `secure = TRUE` in production, where the app is served over https.

Names and values are checked against the character set cookies allow. A value
carrying a `;`, a newline or a quote is rejected rather than silently truncating
the header or, worse, appending one of its own.

## Reading it back

From a page server:

```{r}
server <- function(input, output, session) {
  cookies <- parse_cookie_string(get_cookies())
  print(cookies["SESSION"])
}
```

Index with single brackets: a request carrying no `Cookie` header at all gives
nothing to index into, and `[[` raises "subscript out of bounds" where `[`
returns `NA`.

`get_cookies()` gives the raw `Cookie` header; `parse_cookie_string()` turns it
into a named vector.

## Removing it

```{r}
page(
  href = "/logout",
  ui = tagList(h1("You are logged out")),
  res_handlers = list(
    ~ remove_cookie(.x, "SESSION", path = "/")
  )
)
```

Pass `remove_cookie()` the same `path` and `domain` you gave `set_cookie()`. A
browser only replaces a cookie when the name, the path and the domain all
match; with a different path it files the deletion as a separate cookie and
leaves the original alone. This bites hardest behind a reverse proxy, where the
default path is the mount rather than `/`, so the deletion appears to work
locally and does nothing once deployed.

## A session across pages

The cookie holds an identifier; the data lives server side, keyed by it.

```{r}
store <- cachem::cache_disk()

# 32 hex characters from the system's cryptographic source. `sample()` would
# read from R's default generator, whose state is recoverable from its output:
# fine for a simulation, not for something that has to be unguessable.
new_session_id <- function() {
  paste(
    sprintf("%02x", as.integer(openssl::rand_bytes(16))),
    collapse = ""
  )
}

login <- function() {
  page(
    href = "/login",
    ui = tagList(h1("Welcome back")),
    res_handlers = list(
      # Runs once per request, so each visitor gets an identifier of their own.
      # Building it in `login()` instead would run it once, when the app is
      # assembled, and hand every visitor the same session.
      function(res, req) {
        id <- new_session_id()
        store$set(id, list(user = "colin", at = Sys.time()))
        set_cookie(res, "SESSION", id, path = "/")
      }
    )
  )
}

home <- function() {
  page(
    href = "/",
    ui = tagList(h1("Home"), verbatimTextOutput("who")),
    server = function(input, output, session) {
      output$who <- renderPrint({
        # Single brackets: with no Cookie header at all there is nothing to
        # index into, and `[[` would raise "subscript out of bounds" before
        # the fallback below is reached.
        id <- parse_cookie_string(get_cookies())["SESSION"]
        if (is.na(id)) "anonymous" else store$get(id)
      })
    }
  )
}
```

The store is what decides what a session may do; the cookie only names it. So
the identifier has to be unguessable, and it has to be minted per visitor —
two mistakes that are easy to make and silent when made.
