---
title: "Deploying behind a reverse proxy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deploying behind a reverse proxy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Your app works. Every page, every link, every plot, at
`http://127.0.0.1:3000`. You deploy it, open
`https://example.com/myapp/`, and the home page shows up unstyled — or the
links go nowhere, or a page comes back "Not found".

Nothing is broken. The app simply does not know it has been moved.

## Does this concern you?

Look at the address your users will type.

- `https://example.com/` — the app is at the root of its domain. Nothing to do,
  stop reading.
- `https://example.com/myapp/` — the app is served under a **prefix**. Read on.
- Posit Connect — a prefix too, but brochure works it out on its own. See below.

## Why a prefix is a problem here

In a normal Shiny app the url is decoration. In brochure it is the app: the url
is what decides which page you get.

You wrote:

```{r}
page(href = "/contact", ui = tagList(h1("Contact")))
```

Deployed under `/myapp/`, the browser asks for `/myapp/contact`. That is not
`/contact`, so no page matches, and you get the 404. The other half of the
problem is the reverse: your link `<a href="/contact">` sends the visitor to
`https://example.com/contact`, outside the app entirely.

## The fix

Tell the app where it lives:

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  page(href = "/contact", ui = tagList(h1("Contact"))),
  basepath = "myapp"
)
```

That is all. Your `page()` hrefs stay as they are — `/` and `/contact`, written
as if the app were at the root — and brochure adds and removes the prefix on
the way in and out.

`"myapp"`, `"/myapp"` and `"/myapp/"` all mean the same thing, so you do not
have to remember which one it wanted.

## On Posit Connect

Nothing to do. Connect tells the app where it is mounted, on every request, and
brochure picks it up:

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  page(href = "/contact", ui = tagList(h1("Contact")))
)
```

Deploy that as it is; `/contact` will work under
`https://connect.example.com/content/<guid>/contact`.

You *may* still set `basepath` if you would rather not depend on that. What
Connect sends is not part of a documented, promised interface — it works, it is
tested, but Posit never committed to keeping it. Setting `basepath` costs you
nothing and pins the behaviour.

The detection only runs when the app really is on Connect. Behind your own
nginx, Apache or Traefik, `basepath` is not optional.

## Reading the symptoms

| What you see | What it means |
| --- | --- |
| The page loads but has no styling, and nothing is interactive | The css and javascript are 404ing. Set `basepath`. |
| A link takes you out of the app, or to a 404 | The links are missing the prefix. Set `basepath`. |
| One page works, a deeper one does not | Same cause, and deeper pages fail first. Set `basepath`. |
| Everything works but `server_redirect()` leaves the app | The redirect target is missing the prefix. Set `basepath`. |
| It works locally and fails deployed, with `basepath` set | Check the value: it is the prefix in the url, without the host. |

A single cause, a single fix. If setting `basepath` changes nothing, the value
is probably wrong: for `https://example.com/team/myapp/`, it is `"team/myapp"`,
not `"myapp"`.

## Writing links that survive

Write internal links as absolute paths, the way you wrote your hrefs:

```{r}
tags$a(href = "/contact", "Contact")        # yes
tags$a(href = "contact", "Contact")         # depends on the current page
tags$a(href = "../contact", "Contact")      # depends on the current page
```

Only the first is rewritten with the prefix. The other two are resolved by the
browser against whatever page the visitor happens to be on, which gives a
different answer on `/` than on `/team/profile`.

## Test it deployed

A prefix is the kind of thing that passes every local test and fails the moment
it ships, because at `127.0.0.1:3000` there is no prefix to get wrong. If you
can, deploy once to a staging url with a prefix and click through the app.

---

## What happens under the hood

You do not need this to use `basepath`, but it explains what you are looking at
in the page source.

On the way in, the prefix is removed from the requested path before brochure
matches it against your pages — so `/myapp/contact` becomes `/contact`. Some
proxies strip the prefix themselves before passing the request on, Connect
among them; brochure only removes it when it is actually there, so both kinds
of proxy work.

On the way out, every `src` and `href` in the response is made absolute:
`href="/contact"` becomes `href="/myapp/contact"`. This also fixes something
that has nothing to do with proxies. Shiny declares its own dependencies with
relative urls, such as `shiny-javascript-1.14.0/shiny.min.js`. On a page like
`/user/colin`, a browser resolves that against `/user/` and asks for a file
that is not there — which is why deeper pages break first. So the rewriting
happens on every response, and with no prefix it simply produces
`/shiny-javascript-1.14.0/shiny.min.js`.

Urls pointing at other hosts, protocol relative ones, fragments and query
strings are left untouched.
