---
title: "Deploying behind a reverse proxy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deploying behind a reverse proxy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r}
library(shiny)
library(brochure)
```

A single page Shiny app barely notices where it is served from. A brochure app
does, because it is the URL that decides which page you get. Served at the root
of a domain there is nothing to do; served under a prefix, the app has to know
about it.

## The problem

Say your app is at `https://example.com/myapp/` and you declared
`page(href = "/contact")`. Two things have to line up:

- a request for `/myapp/contact` must match the href `/contact`
- the link `<a href="/contact">` must be emitted as `/myapp/contact`, or it
  will leave the app

## `basepath`

Tell the app where it is mounted, and both are handled:

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  page(href = "/contact", ui = tagList(h1("Contact"))),
  basepath = "myapp"
)
```

The prefix is removed from the incoming path before it is matched, and
prepended to the urls the app emits. `"myapp"`, `"/myapp"` and `"/myapp/"` all
mean the same thing.

This works whether or not the proxy strips the prefix on its way in. Some do,
some do not; brochure only removes the prefix when it is actually there.

## Posit Connect

Connect serves content under `/content/<guid>/` and tells the app so on every
request. Brochure reads that and configures itself, so `basepath` is not
needed:

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  page(href = "/contact", ui = tagList(h1("Contact")))
)
```

The header this relies on is not part of any published Connect contract, and
the detection only runs when the environment says the app really is on Connect
— elsewhere the header would just be something a client sent. If you would
rather pin the behaviour than rely on it, set `basepath` anyway: an explicit
`basepath` and Connect's own mount describe the same place.

## What gets rewritten

Every `src` and `href` in the response is made absolute to the mount, including
the dependencies Shiny emits relative. That is not only a proxy concern: on a
page with more than one segment, such as `/user/colin`, a relative
`shiny-javascript-1.14.0/shiny.min.js` resolves against `/user/` and 404s. So
the rewriting happens for every response, with an empty mount simply yielding
`/shiny-javascript-1.14.0/...`.

Absolute urls to other hosts, protocol relative ones, fragments and query
strings are left alone.

## Redirections

`server_redirect()` goes through the same mount, so a target written the way
you declared your pages arrives where you expect:

```{r}
server <- function(input, output, session) {
  observeEvent(input$go, {
    server_redirect("/contact")
  })
}
```

Under a mount, this navigates to `/myapp/contact`. Targets carrying a scheme
other than `http(s)` are refused, so a value that reached `to` from user input
cannot turn into a `javascript:` url.

## A checklist

- pages declared with hrefs relative to the app, not to the domain
- `basepath` set, unless the platform announces the mount itself
- links written as `/contact`, never `../contact`
- the app tested at the deployed URL, not only at `127.0.0.1`

That last one matters more than it sounds. A prefix is the kind of thing that
works in every local test and fails the moment it is deployed.
