---
title: "Using brochure with golem"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using brochure with golem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`{golem}` organises a Shiny application as a package with one UI and one
server, assembled in `app_ui.R` and `app_server.R` and launched by `run_app()`.
A brochure app has no single UI and no single server: it has one of each *per
page*. The two fit together well, but the shape of the app changes, and this
article is about that shape rather than about the two functions brochure
exports — those are in `?golem_hook` and `?new_page`.

## What changes

| golem, as usual | golem with brochure |
| --- | --- |
| `app_ui.R` assembles the whole interface | gone; each page brings its own UI |
| `app_server.R` holds the whole server | gone; each page brings its own server |
| modules are pieces of one page | a module *is* a page, plus a `page()` around it |
| `run_app()` calls `shinyApp()` | `run_app()` calls `brochureApp()` |

`golem_add_external_resources()` stays, and matters more than before: it is the
one place declaring your css, javascript and favicon, and it has to reach every
page.

## A fresh app

```{r}
golem::create_golem("myapp", project_hook = brochure::golem_hook)
```

The hook removes `app_ui.R` and `app_server.R`, writes a `run_app()` built on
`brochureApp()`, and drops in a first page, `R/mod_home.R`. It also rewrites the
`golem::add_module()` calls in `dev/02_dev.R` so they use the brochure module
template — the file you work from already does the right thing.

## Adding a page

```{r}
golem::add_module(name = "contact", module_template = brochure::new_page)
```

You get a file holding three things: a module UI, a module server, and a
function wrapping both into a `page()`:

```{r}
mod_contact_ui <- function(id) {
  ns <- NS(id)
  tagList(
    h1("Hello {brochure}!")
  )
}

mod_contact_server <- function(id) {
  moduleServer(id, function(input, output, session) {
    ns <- session$ns
  })
}

contact <- function(id = "contact", href = "/contact") {
  page(
    href = href,
    ui = mod_contact_ui(id = id),
    server = function(input, output, session) {
      mod_contact_server(id = id)
    }
  )
}
```

That last function is what you add to `run_app()`. The module keeps its
namespace, so two pages may reuse the same module under different ids and
hrefs — a `mod_profile` mounted once at `/me` and once at `/user/:id`.

## Wiring it up

```{r}
run_app <- function(
  onStart = NULL,
  options = list(),
  enableBookmarking = NULL,
  ...
) {
  with_golem_options(
    app = brochureApp(
      # Injected into every page: css, javascript, favicon
      golem_add_external_resources(),
      # One entry per page
      home(),
      contact(),
      onStart = onStart,
      options = options,
      enableBookmarking = enableBookmarking
    ),
    golem_opts = list(...)
  )
}
```

`golem_add_external_resources()` goes in `...`, among the pages, because
`brochureApp()` injects everything there that is not a page or a redirection
into all of them. Put it inside a single `page()` and only that page gets your
stylesheet.

## Adapting an existing golem app

1. delete `R/app_server.R`, and keep only `golem_add_external_resources()` from
   `R/app_ui.R`
2. turn each screen into a file exposing a `page()`, the way `new_page()` lays
   it out
3. replace `shinyApp()` with `brochureApp()` in `run_app()`, passing
   `golem_add_external_resources()` and then the pages

Step 2 is where the work is, and it is rarely mechanical. Anything the old
`app_server.R` held to pass state between screens has no equivalent: a page
cannot read another page's reactive values, because they never exist at the
same time. That state has to move to a cookie plus a store — see
`vignette("cookies")`.

## What stays golem

`golem::get_golem_options()`, the config file, `app_sys()`, the `dev/` scripts,
`golem::add_dockerfile()`: all of it works unchanged. `run_app()` is still the
entry point, and the app is still a package.

Testing changes a little: there is no single `app_server()` to drive with
`shiny::testServer()`. Test each page's module server as you would any module,
and drive the app's routing through its `httpHandler` with a request built by
hand — that is how brochure tests itself, in
`tests/testthat/test-brochureApp.R`.
