---
title: "Request and response handlers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Request and response handlers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

```{r}
library(shiny)
library(brochure)
```

Between the moment a browser asks for a page and the moment it gets an answer,
brochure gives you two places to intervene. If you know `express.js`, this is
its middleware; in `{shiny}` terms it is code that runs *before* any reactive
context exists.

## The order things run in

1. the browser sends a `GET`, which R receives as a `req` object
2. the app level `req_handlers` run, in order
3. the request is matched against your pages
4. the matched page's `req_handlers` run
5. R builds the `httpResponse` from that page's UI
6. the app level `res_handlers` run, then the matched page's
7. the response is sent

A `req_handler` takes `req` and must return it. A `res_handler` takes `res` and
`req`, and must return `res`. Both accept formulas: `.x` is `req` for a request
handler, and `res` for a response handler, where `.y` is then `req`.

## Logging every request

```{r}
log_where <- function(req) {
  message(Sys.time(), " - ", req$PATH_INFO)
  req
}

brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  req_handlers = list(log_where)
)
```

## Answering before Shiny

A `req_handler` returning an `httpResponse` short-circuits: that response is
sent immediately, nothing else runs, and no Shiny session is created. This is
how you serve something that is not a page — a healthcheck, a webhook receiver,
a small JSON endpoint:

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  page(
    href = "/healthcheck",
    ui = tagList(),
    req_handlers = list(
      ~ shiny::httpResponse(200, content = "OK")
    )
  )
)
```

The page still needs a `ui`, because `page()` requires an href to bind to, but
it is never rendered.

Note the asymmetry: an early `httpResponse` skips the `res_handlers` too. If
you rely on a response handler to set a header on every answer, it will not see
these.

## Rejecting a request

Because a request handler can answer on its own, it can also refuse:

```{r}
require_token <- function(req) {
  token <- req$HTTP_X_TOKEN
  if (is.null(token) || !identical(token, Sys.getenv("APP_TOKEN"))) {
    return(shiny::httpResponse(401, content = "Unauthorized"))
  }
  req
}

brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  req_handlers = list(require_token)
)
```

This guards the HTTP request that serves the page. It does not guard the
websocket the page opens afterwards, so treat it as a gate on the document, not
as an authentication system.

## Changing the response

```{r}
brochureApp(
  page(href = "/", ui = tagList(h1("Home"))),
  res_handlers = list(
    function(res, req) {
      res$headers$`X-Frame-Options` <- "DENY"
      res
    }
  )
)
```

App level handlers run before page level ones, so a page can override what the
app decided.

## Page level or app level

An app level handler runs for every matched page. A page level one runs only
for its own page, which is what you want for anything specific — a cookie for
the login page, a `Cache-Control` for a page that is expensive to build.

Neither runs for a request that matches no page: those get `content_404`
directly.
